---
title: "Practice"
subtitle: "Workshop Best Practices for Reproducible Data Processing and Analysis using R and Quarto"
author: "S.A. & S.P."
date: today
format:
  html:
    theme:
      light: flatly
      dark:  darkly
    fontsize: .9em
    page-layout: full
    toc: true
    toc-depth: 4
    toc-location: left
    toc-title: Contents
    toc_float:
       collapsed: false
    number-sections: false
    number-depth: 4
    grid:
      sidebar-width: 150px
      body-width: 1500px
      margin-width: 50px
      gutter-width: 1.5rem
    smooth-scroll: true
    highlight-style: github
    date-format: "D MMMM YYYY"
    code-fold: false
    code-summary: "Show the code"
    code-overflow: wrap
    code-tools: true

execute:
  echo: true
  warning: false
  output: true
  error: false
embed-resources: true
fig-responsive: false
standalone: true
---

Writing robust code is essential, especially when working with complex data, collaborating with others, or ensuring long-term reproducibility.

# Setting Up

```{r}
#| output: false

# 1. 
# For this code block, it would be preferable not to print the output, 
# just show the code. How can we configure that?


# 2.
# We want to set the working directory to our folder automatically.
# How can we do that without hardcoding the path?
# Don't forget to comment out this line of code to ensure proper rendering.

#...............................................................................
#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))                 
#...............................................................................


# 3.
# Now, we want to install the libraries, but only the missing ones.
# We also want to load all the libraries we will use in this script.
# How we can do that without repeating the same code?
# We need the package qualtRics, readr, dplyr, ggplot2, janitor, lme4, 
# kableExtra, report, visdat

# ...............................................................................
# Load packages
list.of.packages <- c(
  "qualtRics",         # Read files obtained through Qualtrics
  "readr",             # Write csv files
  "dplyr",             # Manipulate data during preprocessing
  "janitor",           # Clean and manage data during preprocessing
  "ggplot2",           # Create plots
  "kableExtra",        # Display tables in HTML format
  "report",            # Generate advanced reporting for linear models
  "visdat",            # Visualize missing data
  "plotly"             # Generate interactive plots
)

new.packages <-
  list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
if (length(new.packages))
  # Install any new packages
  install.packages(new.packages)
# Library in all packages
lapply(list.of.packages, require, character.only = TRUE)
rm(list.of.packages, new.packages)
# ...............................................................................


# 4.
# We want to disable scientific notation for easier readability of numbers.
# What can we do?

#................................................................................
options(scipen = 999)
#................................................................................


# We advice you to keep these lines of code in your script:

# Set the console output width to improve the readability of printed data.
options("width" = 900)

# Avoid displaying 'NA' values in tables generated with knitr::kable.
options(knitr.kable.NA = '')

```


## Load Data

```{r}
#| output: false

# 5. 
# For this code block, it would be preferable not to print the output, 
# just show the code. How can we configure that?


# If you prefer to use .csv, qualtRics package can be very useful
# to read the data.

# Load the data from Lab UK
raw_data_uk <- qualtRics::read_survey("practice_data_uk.csv")

# Load the data from Lab US
raw_data_us <- qualtRics::read_survey("practice_data_us.csv")


# 6.
# These dataframes are Qualtrics copies of the same survey, 
# so they should have the same columns. But do they?

#...............................................................................
ncol(raw_data_uk)
ncol(raw_data_us)
#...............................................................................

```


## Merge Data

```{r}

# Merge the data from the two labs.
data_merged <- dplyr::full_join(raw_data_uk, raw_data_us)

# 7.
# Let's check if the data was merged correctly. What can we do?
# (don't forget to remove the unnecessary variables after the check)

#...............................................................................
# Row sum of the individual dataframes
nrow(raw_data_uk) + nrow(raw_data_us)

# If the value below is the same as the row sum, the sanity check passed.
nrow(data_merged)

# Columns' names from the individual dataframes merged 
# to compare with columns' names from data_merged
data_merged_col <- names(data_merged)
all_col <- unique(c(names(raw_data_uk),
                        names(raw_data_us)))

# Columns missing in data_merged. If empty, the sanity check passed.
setdiff(all_col, data_merged_col)

# Columns in excess in data_merged. If empty, the sanity check passed.
setdiff(data_merged_col, all_col)

rm("all_col", "data_merged_col")
#...............................................................................

```


## Ethical Compliance Check

```{r}

# The minimum age to participate in the survey was 18 years old.  
# We need to check for underage participants in our data 
# and wipe out their responses except their ResponseID.

min(data_merged$age)


# We need to verify if any participant wrote hateful content or identifiable  
# information in the text entries of the survey. If so, we need to remove those 
# participants.

unique(data_merged$sex_55_TEXT)


```


## Save Public Data

```{r}

# 8.
# Now that we have verified the raw data is ready to be shared,  
# we need to save it as a public data version. How can we do that?

#...............................................................................
write_csv(data_merged, "01_public_data.csv")
save(data_merged, file = "01_public_data.Rda")
#...............................................................................

# Let the reader have a overview of the data structure
str(data_merged)

```


# Create a Codebook

```{r}

# If we are going to publish the public data,  
# we should generate the codebook with the current variable names.

codebook <- qualtRics::extract_colmap(raw_data_uk) %>%
  # Keep only relevant columns
  dplyr::select(qname, description) %>%
  # Rename for clarity
  dplyr::rename(variable = qname) 

write.csv(codebook, "codebook_structure.csv")

rm(codebook)
```


# Data Preprocessing

```{r}

data_cleaned <- data_merged


# 9.
# Let's clean the data before we start the analysis.
# We should standardize the column names (e.g., `Duration (in seconds)`) to
# follow the snake_case convention. How can we do that?

#...............................................................................
data_cleaned <- janitor::clean_names(data_cleaned)
#...............................................................................



# 10.
# All items had forced response feature activated. But are we sure?

#...............................................................................
vis_miss(data_cleaned)
#...............................................................................



# 11.
# Now let's check the variable age.
# Is there something wrong with it?

mean(data_cleaned$age)

#...............................................................................
# Always check the structure of the variable.
# There are 11 answers reporting an age of 366 years. 
# Something is wrong, probably the recoding of the options.

table(data_cleaned$age)

# Let's recode the variable age to the correct values.
data_cleaned$age <- ifelse(data_cleaned$age == 366, 36, data_cleaned$age)
#...............................................................................



# 12.
# Let's reverse the "pray" variable. 
# Currently, the scale is 1 = Several times a days
# to 8 = Never, practically never.
# How can we conduct this operation in a robust way?

#...............................................................................
table(data_cleaned$pray)

data_cleaned$pray <- 9 - data_cleaned$pray

table(data_cleaned$pray)
#...............................................................................


# 13.
# We want to create an index by averaging the FSES items for each participant. 
# How can we do this in a robust way?

#...............................................................................
data_cleaned$fses_avg <- rowMeans(data_cleaned[, c("fses1", "fses2", "fses3", "fses4", "fses5", "fses6")], na.rm = TRUE)

# Manually compute the index for one participant and compare it
(data_cleaned$fses1[1] + data_cleaned$fses2[1] + data_cleaned$fses3[1] + data_cleaned$fses4[1] + data_cleaned$fses5[1] + data_cleaned$fses6[1])/6

data_cleaned$fses_avg[1]

#...............................................................................




```


# Save the Cleaned Data

```{r}

# 14.
# Now that we have cleaned the raw data,  
# we need to save it as a cleaned data version. How can we do that?

#...............................................................................
write_csv(data_cleaned, "02_data_cleaned.csv")
save(data_cleaned, file = "02_data_cleaned.Rda")
#...............................................................................

```


# Data Visualization

```{r}

# 15.
# We want to create a tabset panel with two separators and fold the code.
# How can we do that? 

```

## Financial Self-Efficacy

::: {.panel-tabset}

### by Sex

```{r}
#| code-fold: TRUE

# Filter data to keep only sex values 0 and 1
data_filtered <- subset(data_cleaned, sex %in% c(0, 1))

# Convert sex to a factor with labels
data_filtered$sex_label <- factor(data_filtered$sex,
                                  levels = c(0, 1),
                                  labels = c("Male", "Female"))

# Create the ggplot
p_fses_sex <- ggplot2::ggplot(data_filtered, aes(x = sex_label, 
                                        y = fses_avg, 
                                        fill = sex_label)) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_boxplot(width = 0.1, fill = "white", outlier.shape = NA) +
  labs(x = "Sex", 
       y = "Financial Self-Efficacy", 
       title = "Financial Self-Efficacy by Sex") +
  theme_minimal() +
  scale_fill_manual(values = c("#43CD80", "#EE6363")) +
  theme(legend.position = "none")

# Convert to interactive ggplotly
plotly::ggplotly(p_fses_sex)

```

### by Age
```{r}
#| code-fold: TRUE

# Create the ggplot
p_fses_age <- ggplot2::ggplot(data_filtered, aes(x = age, y = fses_avg)) +
  geom_point(alpha = 0.5, color = "#00BFFF") + 
  geom_smooth(method = "lm", color = "#0000CD", se = FALSE) +
  labs(x = "Age", 
       y = "Financial Self-Efficacy", 
       title = "Financial Self-Efficacy by Age") +
  theme_minimal()

# Convert to interactive ggplotly
plotly::ggplotly(p_fses_age)

```

:::


# Data Analysis

## Model: Financial Self-Efficacy by Sex
```{r}

m_fses_sex <- stats::lm(fses_avg ~ sex, data = data_filtered)

# Report the results
knitr::kable(report_table(m_fses_sex))

# Report the effect size
effectsize::cohens_d(fses_avg ~ sex, data = data_filtered)

```

### Check the Assumptions
```{r}

m_fses_sex_fit <- performance::check_model(m_fses_sex)
m_fses_sex_fit

```


# Information About the R Session

```{r}

sessionInfo()

```


# Resources for you to explore later

## Configure Qualtrics API to extract data directly as Rda file
```{r}
#| eval: false

# You only need to run this code once on your computer to set it up permanently.
# There is no need to include this code in all scripts.
# Do not share your API token with anyone.

qualtrics_api_credentials(
  # check your Qualtrics account settings to get your API key
  api_key = "<YOUR-QUALTRICS_API_KEY>", 
  # replace by your university base url
  base_url = "https://vuamsterdam.eu.qualtrics.com/",
  install = TRUE)


# Then you only need the following code to access your surveys

# Load the data from Qualtrics
raw_data <- fetch_survey(
    # replace by your survey ID "SV_XXXXXX". 
    # you can see your survey ID in the data collection link.
    surveyID = "<YOUR-SURVEY-ID>",
    # if TRUE it will return text of answer choices, instead of recoded values.
    label = FALSE,
    convert = FALSE,
    include_display_order = FALSE)

```

